---
title: "Analysis of AD polygenic risk scores (AD-PRS) with cytokine expression"
author: |
 | Juliana Beker Godoy, MSc
 | Ricardo A. Vialle, PhD
 | Katia de Paiva Lopes, PhD
 | Graduate Program in Bioinformatics
 | Professional and Technological Education Sector
 | Federal University of Parana (UFPR)
 | Curitiba, Brazil
date: "`r Sys.Date()`"
output:
 rmarkdown::html_document:
   theme: cerulean
   highlight: haddock
   code_folding: hide
   toc: true
   toc_float: true
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

> single-nuclei RNAseq dataset | 424 unique samples from DLPFC

```{r clean, warning=FALSE, echo=FALSE, results='hide'}
#This command clean all variables. BE CAREFULL!!! 
rm(list = setdiff(ls(), lsf.str()))
```

```{r HelperFunctions, echo = FALSE}
# Function to create useful tables
createDT <- function(DF, caption="", scrollY=500){
  data <- DT::datatable(DF, caption=caption,
   extensions =  'Buttons',
   options = list( dom = 'Bfrtip', 
                    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'), 
                    scrollY = scrollY, scrollX=T, scrollCollapse = T, paging = F,  
                    columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
   ) 
   return(data)
}
```

```{r packages, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(data.table)
library(ggeasy)
library(ggpubr)
library(ggbeeswarm)
library(scales)
library(ggsci)
library(patchwork)
library(gtools)
library(circlize)
library(RColorBrewer)
library(readxl)
library(fgsea)
library(foreach)
library(doParallel)
library(openxlsx)
```

```{r set_dir, warning=FALSE, message=FALSE, include=FALSE} 
work_dir <- "/pastel/projects/cytokines_Juliana/JNI_revision_I/"
```

# PRS analysis

## Read data and prepare inputs

```{r , warning=FALSE, message=FALSE, include=FALSE} 
# Expression data - batch adjusted
celltype_exp <- readRDS(file.path(work_dir, "snRNA_pseudoBulk_7majCellTypes.ComBat.rds"))
maj_celltypes = names(celltype_exp)

# PRS data
prs_combined = fread(paste0("/pastel/projects/cytokines_Juliana/PRS_analysis/PRS_combined.csv")) %>% as.data.frame()

prs_id = "ADPRS_Bellenguez_2022_Pt_0.001"
prs = prs_combined[,c("projid",prs_id)]
colnames(prs) = c("projid","prs")

# Getting only the volunteers (projid) present on single nuclei data (n=424)
donors_in_common = intersect(colnames(celltype_exp$ext),prs$projid)
length(donors_in_common)

ancestry_PCs = read_csv("/pastel/projects/sv_and_resilience/pca.ancestry_pcs.csv")
idmapping = read_delim(paste0("/pastel/projects/sv_and_resilience/ROSMAP_IDkey_with_fixed_projids.csv"),
                       guess_max = 1e6, show_col_types = FALSE)
# Map sample ids in the genetic PCs file
ancestry_PCs = ancestry_PCs %>% left_join(unique(idmapping[,c("projid","wgs_id")]), by = c("ancestry"="wgs_id")) %>% as.data.frame()
ancestry_PCs$ancestry = ifelse(!is.na(ancestry_PCs$projid), "ROSMAP", ancestry_PCs$ancestry)

prs = prs %>%
  left_join(ancestry_PCs, by = "projid") %>%
  filter(projid %in% donors_in_common) %>% na.omit() %>%
  mutate(prs_residuals = resid(lm(prs ~ PC1 + PC2 + PC3, data = .))) %>%
  select(projid, prs_residuals) %>% distinct() %>% mutate(prs = prs_residuals)

pheno_SN = as.data.frame(readRDS(file.path(work_dir, "phenotypes_updated.rds")))
rownames(pheno_SN) = pheno_SN$projid
pheno_SN = pheno_SN %>% 
  left_join(prs) 

# Replace NA values in 'pmi' with the mean of non-NA values
pheno_SN$pmi[is.na(pheno_SN$pmi)] <- mean(pheno_SN$pmi, na.rm = TRUE)

# Cytokines list
cytokines_list = read_csv("/pastel/resources/data2share/Cytokines with families - Fewer groups.csv", show_col_types = FALSE)
cytokines_list$...1 = NULL
cytokines <- read.delim( paste0(work_dir, "list_cytokines_T015.txt"), header = TRUE, check.names = FALSE, stringsAsFactors = FALSE)
cytokines <- subset(cytokines, select = -family) 
ensembls = cytokines$ensembl
```

## Results from linear/logistic regression

```{r}
load(file.path(work_dir, "cytokines_snRNA_APOE_association_results.RData")) # assoc_results
assoc_results_APOE = assoc_results

load(file.path(work_dir, "cytokines_snRNA_PRS_association_results.RData")) # assoc_results
assoc_results_ADPRS = assoc_results
```

<!----------- Start -------------->

## PRS deciles distribution

```{r prs, warning=FALSE, message=FALSE, fig.width=4, fig.height=3} 
# New df containing only interested PRS values
prs = prs_combined[,c("projid",prs_id)]
colnames(prs) = c("projid","prs")

# Getting only the volunteers (projid) present on single nuclei data (n=424)
donors_in_common = intersect(colnames(celltype_exp$ext),prs$projid)

# Spliting data into deciles
PRS_quantiles = prs %>% 
  filter(projid %in% donors_in_common) %>%
  mutate(prs_quantile = as.factor(cut(prs,quantile(prs, probs = 0:10 / 10, na.rm = T),include.lowest = T, labels = F)))
# table(PRS_quantiles$prs_quantile, useNA = "ifany")

# Show PRS distribution in each of the ten quantile
(p1 = ggplot(PRS_quantiles, aes(x = prs, fill = prs_quantile)) +
    geom_histogram(alpha = 1, bins = 100) +
    theme_classic() +
    scale_fill_viridis_d() +
    labs(x = "PRS", y = "Density") +
    theme(
        legend.position = "none"
    ))

min_quantile = min(as.numeric(PRS_quantiles$prs_quantile), na.rm = T)
max_quantile = max(as.numeric(PRS_quantiles$prs_quantile), na.rm = T)

# Make a long table of expression of cytokines in each cell type and PRS group
celltype_exp_prs_long = data.frame()
for(cell_i in maj_celltypes){
  celltype_exp_prs_long = bind_rows(celltype_exp_prs_long,
                              celltype_exp[[cell_i]] %>% as.data.frame() %>% rownames_to_column("ensembl") %>%
                                filter(ensembl %in% cytokines_list$ensembl) %>%
                                pivot_longer(cols = -ensembl, names_to = "projid", values_to = "expression") %>%
                                mutate(celltype = cell_i) %>%
                                left_join(PRS_quantiles, by = "projid") %>% na.omit())
}

celltype_exp_prs_long = celltype_exp_prs_long %>% left_join(cytokines_list, by = "ensembl")
```

## Top cytokines correlated with PRS/APOE

Top 5 cytokines significantly associated with AD-PRS calculated from GWAS study published by Bellenguez et al. (2022). This article can be accessed [here](https://www.nature.com/articles/s41588-022-01024-z)
And Top 5 cytokines significantly associated with APOE4 genotype.

```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=3}
# Plot scatter plots for the top5 cytokines correlated with PRS
assoc_results_ADPRS_all = lapply(assoc_results_ADPRS, function(x) x$all_stats_df) %>% 
  bind_rows(.id = "cell_type") %>% 
  mutate(cell_type = factor(cell_type, levels = names(celltype_exp))) %>%
  mutate(cell_symbol = paste(cell_type, predictor, sep = ":")) %>%
  arrange(fdr)

assoc_results_ADPRS_all <- assoc_results_ADPRS_all %>%
  mutate(cell_type = case_when(
    cell_type == "ext" ~ "Excitatory Neurons",
    cell_type == "inh" ~ "Inhibitory Neurons",
    cell_type == "oli" ~ "Oligodendrocytes",
    cell_type == "mic" ~ "Microglia",
    cell_type == "ast" ~ "Astrocytes",
    cell_type == "end" ~ "Endothelial Cells",
    cell_type == "opc" ~ "OPCs",
    TRUE ~ cell_type  # keep original value if none match
  ))

top_cytokines_PRS = assoc_results_ADPRS_all %>% ungroup() %>% slice_min(order_by = p, n = 5) 

# Plot scatter plots for the top5 cytokines correlated with PRS
assoc_results_APOE_all = lapply(assoc_results_APOE, function(x) x$all_stats_df) %>% 
  bind_rows(.id = "cell_type") %>% 
  mutate(cell_type = factor(cell_type, levels = names(celltype_exp))) %>%
  mutate(cell_symbol = paste(cell_type, predictor, sep = ":")) %>%
  arrange(fdr)

assoc_results_APOE_all <- assoc_results_APOE_all %>%
  mutate(cell_type = case_when(
    cell_type == "ext" ~ "Excitatory Neurons",
    cell_type == "inh" ~ "Inhibitory Neurons",
    cell_type == "oli" ~ "Oligodendrocytes",
    cell_type == "mic" ~ "Microglia",
    cell_type == "ast" ~ "Astrocytes",
    cell_type == "end" ~ "Endothelial Cells",
    cell_type == "opc" ~ "OPCs",
    TRUE ~ cell_type  # keep original value if none match
  ))

top_cytokines_APOE = assoc_results_APOE_all %>% ungroup() %>% slice_min(order_by = p, n = 5) 

top_cytokines = bind_rows(top_cytokines_PRS, top_cytokines_APOE) 
```

```{r top5_cytokine_prs_association, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
top_cytokines_data_plus_with_stats = celltype_exp_prs_long %>%
    mutate(cell_symbol = paste(celltype, symbol, sep = ":")) %>%
    filter(cell_symbol %in% top_cytokines$cell_symbol) %>%
    left_join(assoc_results_ADPRS_all, by = "cell_symbol") %>%
    mutate(label = paste("nom.P =", format(p, digits=1, nsmall=3)))
top_cytokines_data_plus_with_stats$predictor = factor(top_cytokines_data_plus_with_stats$predictor,
                                                      levels = unique(top_cytokines$predictor))

(p2 = ggplot(top_cytokines_data_plus_with_stats, aes(x = prs, y = expression, color = cell_type)) +
    geom_point(alpha = 0.3) + 
    geom_smooth(method = "lm", se = T) + 
    geom_text(aes(x = Inf, y = -Inf, label = label),
              hjust = 1.1, vjust = -0.8, size = 4.5, color = "black", check_overlap = T) +
    facet_wrap(symbol ~ cell_type, scales = "free", ncol = 5, nrow = 2) +
    theme_classic() + 
    scale_x_continuous(n.breaks = 3) +
    # ggeasy::easy_rotate_x_labels(angle = 25, side = c("right")) +
    theme(legend.position = "none") +
    labs(x = "PRS", y = "Expression") +
    theme(
      legend.position = "none"
    ) +
    scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    ))
)
```

## Top cytokines correlated with APOE

```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
top_cytokines_APOE_data_plus_with_stats = celltype_exp_prs_long %>%
    mutate(cell_symbol = paste(celltype, symbol, sep = ":")) %>%
    filter(cell_symbol %in% top_cytokines$cell_symbol) %>%
    left_join(assoc_results_APOE_all, by = "cell_symbol") %>%
    left_join(pheno_SN[,c("projid","apoe_any4")]) %>%
    mutate(label = paste("nom.P =", format(p, digits=1, nsmall=3)))
top_cytokines_APOE_data_plus_with_stats$predictor = factor(top_cytokines_APOE_data_plus_with_stats$predictor,
                                                      levels = unique(top_cytokines$predictor))

(p3 = ggplot(na.omit(top_cytokines_APOE_data_plus_with_stats), 
       aes(x = apoe_any4, y = expression, color = cell_type)) + 
  geom_boxplot(notch=F, fill="white", color="black", outlier.shape=NA, lwd=0.4, alpha=1, width=.7, fatten = 3) +
  ggbeeswarm::geom_quasirandom(alpha=.3, width=.2, show.legend=F, size=2) + 
  geom_pwc(method = "t_test", label.y.npc = "top", label.size = 5, hide.ns = F) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
                   geom = "crossbar", width = .7, color = "black") +
  geom_text(aes(x = Inf, y = Inf, label = label),
              hjust = 1.1, vjust = 1.2, size = 4.5, color = "black", check_overlap = T) +
  facet_wrap( ~ symbol + cell_type, scales = "free", nrow = 2, ncol = 5) +
  # facet_wrap(~ celltype+symbol, scales = "free_y") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
      scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    )) +
  xlab("APOE e4") + 
  ylab("Expression") +
  theme_classic() +
  theme(legend.position = "none"))
```

## Permutation test

```{r warning=FALSE, message=FALSE, eval=T}
library(dplyr)
library(furrr)
library(data.table)
library(parallel)
library(future)
library(future.apply)

nperm = 1000
set.seed(123)
seeds <- sample(1:10000, nperm)
# Permute cell-type labels within each individual while keeping PRS and expression values fixed
permute_prs <- function(celltype_exp_prs_long, seeds, n_cores = NULL) {
  # Set up parallel processing
  if (is.null(n_cores)) {
    n_cores <- min(detectCores() - 1, length(seeds))
  }
  # Convert to data.table for faster operations
  dt <- as.data.table(celltype_exp_prs_long)
  dt[, ...1 := NULL]  # Remove unnecessary column
  
  # Define function for each permutation using data.table
  process_permutation_dt <- function(seed_val, data) {
    set.seed(seed_val)
    
    # Create copy and permute
    dt_perm <- copy(data)
    dt_perm[, prs_perm := sample(prs), by = ensembl]
    
    # Calculate correlations
    perm_cor_dt <- dt_perm[, .(
      correlation = cor(prs_perm, expression, method = "spearman", use = "pairwise.complete.obs")
    ), by = .(ensembl, celltype)]
    
    perm_cor_dt[, permutation := seed_val]
    return(perm_cor_dt)
  }
  
  # Use parallel processing with data.table
  plan(multisession, workers = n_cores)
  result_list <- future_map(seeds, ~process_permutation_dt(.x, dt), .options = furrr_options(seed = TRUE))
  plan(sequential)
  
  # Combine results
  rbindlist(result_list)
}

prs_correlation_prs_permuted = permute_prs(celltype_exp_prs_long, seeds)
avg_prs_correlation_prs_permuted = prs_correlation_prs_permuted[, .(avg_correlation = mean(correlation)), by = .(celltype, permutation)]

prs_correlation_celltype_observed <- celltype_exp_prs_long %>%
  group_by(celltype,ensembl) %>%
  summarise(observed_correlation = cor(prs, expression, method = "spearman", use = "pairwise.complete.obs"), .groups = 'drop')

avg_prs_correlation_celltype_observed = prs_correlation_celltype_observed %>% group_by(celltype) %>%
  reframe(avg_observed_correlation = mean(observed_correlation))

avg_pval_df <- as.data.frame(avg_prs_correlation_prs_permuted) %>% 
  left_join(avg_prs_correlation_celltype_observed) %>%
  group_by(celltype) %>%
  summarise(pval = sum(avg_correlation >= avg_observed_correlation) / n()) %>%
  mutate(cell_type = case_when(
    celltype == "ext" ~ "Excitatory Neurons",
    celltype == "inh" ~ "Inhibitory Neurons",
    celltype == "oli" ~ "Oligodendrocytes",
    celltype == "mic" ~ "Microglia",
    celltype == "ast" ~ "Astrocytes",
    celltype == "end" ~ "Endothelial Cells",
    celltype == "opc" ~ "OPCs",
    TRUE ~ celltype  # keep original value if none match
  ))

maj_cell_avg_correlations = avg_prs_correlation_celltype_observed %>% mutate(permutation = "observed") %>%
  rename(avg_correlation = avg_observed_correlation) %>%
  bind_rows(avg_prs_correlation_prs_permuted %>% mutate(permutation = "permuted")) %>%
  left_join(avg_pval_df, by = "celltype") 
```

```{r warning=FALSE, message=FALSE, fig.width=15, fig.height=2.5, eval=T}
# Plot mean permuted distributions vs observed correlation of PRS vs cytokine expression (facet by celltype)
(p4 = maj_cell_avg_correlations %>% filter(permutation != "observed") %>% 
    ggplot(aes(avg_correlation)) +
    geom_histogram() +
    geom_vline(data = maj_cell_avg_correlations %>% filter(permutation == "observed"),
               aes(xintercept = avg_correlation), color = "red") +
    geom_text(data = avg_pval_df,
              aes(x = -Inf, y = Inf,
                  label = paste0("P = ",format(pval, digits=1, nsmall=3))), hjust = -0.1, vjust = 1.5, size = 3.5) +
    facet_wrap(~cell_type, scales = "free_y", nrow = 1) +
    scale_x_continuous(n.breaks = 3) +
    theme_classic()+
    # ggeasy::easy_rotate_x_labels(angle = 25, side = c("right")) +
    labs(x = "Average correlation", y = "Distribution")
)
```

```{r warning=FALSE, message=FALSE, fig.width=10, fig.height=3}
assoc_results_all = bind_rows(assoc_results_ADPRS_all,assoc_results_APOE_all)

assoc_results_all$predictor <- factor(
  assoc_results_all$predictor,
  levels = assoc_results_all %>%
    arrange(outcome, cell_type, predictor) %>% 
    pull(predictor) %>%
    unique()
)

assoc_results_all$outcome[assoc_results_all$outcome == "prs"] <- "AD PRS"
assoc_results_all$outcome[assoc_results_all$outcome == "apoe_any4"] <- "APOE ε4"

# Define the color order for cell types
cell_type_order <- c("Excitatory Neurons", "Inhibitory Neurons", "Microglia", 
                     "Astrocytes", "Oligodendrocytes", "OPCs", "Endothelial Cells")

# Reorder the data by cell_type according to the color order
assoc_results_all$cell_type <- factor(assoc_results_all$cell_type, levels = cell_type_order)

(p_assoc <- ggplot(assoc_results_all, aes(fill=cell_type, x=cell_type, y=log10_p)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
  geom_hline(yintercept = max(assoc_results_all$log10_p[assoc_results_all$fdr>0.05]), linetype="dashed") + 
  theme_classic() +
  easy_remove_x_axis(what = "text") +
  easy_labs(x = "Cytokines", 
            subtitle = "Cytokine expression associated with AD PRS and APOE ε4",
            fill = "Cell Type", 
            y = bquote(-log[10](italic(P)-value))) + 
  geom_text(data = subset(assoc_results_all, fdr<=0.05),
            aes(label = predictor), 
            position=position_dodge(width=0.9), vjust=-0.25, size = 3.5, fontface = "italic") +
  facet_wrap(~outcome, scales = "free", ncol = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    "Excitatory Neurons" = "#00008B",
    "Inhibitory Neurons" = "#CC3311",
    "Microglia" = "#FF8C00",
    "Astrocytes" = "#9932CC",
    "Oligodendrocytes" = "#698b22",
    "OPCs" = "#008b45",
    "Endothelial Cells" = "#8a6407"
  )))

cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I/PRS_APOE_barplot.pdf", width = 10, height = 3)
p_assoc
dev.off()
```

```{r warning=FALSE, message=FALSE, fig.width=5, fig.height=3}
p_SPP1_PRS = top_cytokines_data_plus_with_stats %>% filter(cell_symbol == "mic:SPP1") %>%
   ggplot(aes(x = prs, y = expression, color = cell_type)) +
    geom_point(alpha = 0.3) + 
    geom_smooth(method = "lm", se = T, color = "black") + 
    geom_text(aes(x = Inf, y = -Inf, label = label),
              hjust = 1.1, vjust = -0.8, size = 4.5, color = "black", check_overlap = T) +
    facet_wrap(symbol ~ cell_type, scales = "free", ncol = 5, nrow = 2) +
    theme_classic() + 
    scale_x_continuous(n.breaks = 3) +
    # ggeasy::easy_rotate_x_labels(angle = 25, side = c("right")) +
    theme(legend.position = "none") +
    labs(x = "PRS", y = "Expression") +
    theme(
      legend.position = "none"
    ) +
    scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    ))


p_SPP1_APOE = na.omit(top_cytokines_APOE_data_plus_with_stats) %>%filter(cell_symbol == "mic:SPP1") %>%
    ggplot(aes(x = apoe_any4, y = expression, color = cell_type)) + 
  geom_boxplot(notch=F, fill="white", color="black", outlier.shape=NA, lwd=0.4, alpha=1, width=.7, fatten = 3) +
  ggbeeswarm::geom_quasirandom(alpha=.3, width=.2, show.legend=F, size=2) + 
  geom_pwc(method = "t_test", label.y.npc = "top", label.size = 5, hide.ns = F) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
                   geom = "crossbar", width = .7, color = "black") +
  geom_text(aes(x = Inf, y = Inf, label = label),
              hjust = 1.1, vjust = 1.2, size = 4.5, color = "black", check_overlap = T) +
  facet_wrap( ~ symbol + cell_type, scales = "free", nrow = 2, ncol = 5) +
  # facet_wrap(~ celltype+symbol, scales = "free_y") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
      scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    )) +
  xlab("APOE e4") + 
  ylab("Expression") +
  theme_classic() +
  theme(legend.position = "none")

(SPP1_plot = ggarrange(p_SPP1_PRS, p_SPP1_APOE, ncol = 2, nrow = 1))

cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I//PRS_SPP1_plot.pdf", width = 5, height = 3)
SPP1_plot
dev.off()
```

```{r warning=FALSE, message=FALSE, fig.width=5, fig.height=3}
p_IL15_PRS = top_cytokines_data_plus_with_stats %>% filter(cell_symbol == "mic:IL15") %>%
   ggplot(aes(x = prs, y = expression, color = cell_type)) +
    geom_point(alpha = 0.3) + 
    geom_smooth(method = "lm", se = T, color = "black") + 
    geom_text(aes(x = Inf, y = -Inf, label = label),
              hjust = 1.1, vjust = -0.8, size = 4.5, color = "black", check_overlap = T) +
    facet_wrap(symbol ~ cell_type, scales = "free", ncol = 5, nrow = 2) +
    theme_classic() + 
    scale_x_continuous(n.breaks = 3) +
    # ggeasy::easy_rotate_x_labels(angle = 25, side = c("right")) +
    theme(legend.position = "none") +
    labs(x = "PRS", y = "Expression") +
    theme(
      legend.position = "none"
    ) +
    scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    ))


p_IL15_APOE = na.omit(top_cytokines_APOE_data_plus_with_stats) %>%filter(cell_symbol == "mic:IL15") %>%
    ggplot(aes(x = apoe_any4, y = expression, color = cell_type)) + 
  geom_boxplot(notch=F, fill="white", color="black", outlier.shape=NA, lwd=0.4, alpha=1, width=.7, fatten = 3) +
  ggbeeswarm::geom_quasirandom(alpha=.3, width=.2, show.legend=F, size=2) + 
  geom_pwc(method = "t_test", label.y.npc = "top", label.size = 5, hide.ns = F) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
                   geom = "crossbar", width = .7, color = "black") +
  geom_text(aes(x = Inf, y = Inf, label = label),
              hjust = 1.1, vjust = 1.2, size = 4.5, color = "black", check_overlap = T) +
  facet_wrap( ~ symbol + cell_type, scales = "free", nrow = 2, ncol = 5) +
  # facet_wrap(~ celltype+symbol, scales = "free_y") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
      scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    )) +
  xlab("APOE e4") + 
  ylab("Expression") +
  theme_classic() +
  theme(legend.position = "none")

(IL15_plot = ggarrange(p_IL15_PRS, p_IL15_APOE, ncol = 2, nrow = 1))

cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I//PRS_IL15_plot.pdf", width = 5, height = 3)
IL15_plot
dev.off()
```

```{r warning=FALSE, message=FALSE, fig.width=5, fig.height=3}
p_IL17D_PRS = top_cytokines_data_plus_with_stats %>% filter(cell_symbol == "ast:IL17D") %>%
   ggplot(aes(x = prs, y = expression, color = cell_type)) +
    geom_point(alpha = 0.3) + 
    geom_smooth(method = "lm", se = T, color = "black") + 
    geom_text(aes(x = Inf, y = -Inf, label = label),
              hjust = 1.1, vjust = -0.8, size = 4.5, color = "black", check_overlap = T) +
    facet_wrap(symbol ~ cell_type, scales = "free", ncol = 5, nrow = 2) +
    theme_classic() + 
    scale_x_continuous(n.breaks = 3) +
    # ggeasy::easy_rotate_x_labels(angle = 25, side = c("right")) +
    theme(legend.position = "none") +
    labs(x = "PRS", y = "Expression") +
    theme(
      legend.position = "none"
    ) +
    scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    ))


p_IL17D_APOE = na.omit(top_cytokines_APOE_data_plus_with_stats) %>%filter(cell_symbol == "ast:IL17D") %>%
    ggplot(aes(x = apoe_any4, y = expression, color = cell_type)) + 
  geom_boxplot(notch=F, fill="white", color="black", outlier.shape=NA, lwd=0.4, alpha=1, width=.7, fatten = 3) +
  ggbeeswarm::geom_quasirandom(alpha=.3, width=.2, show.legend=F, size=2) + 
  geom_pwc(method = "t_test", label.y.npc = "top", label.size = 5, hide.ns = F) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
                   geom = "crossbar", width = .7, color = "black") +
  geom_text(aes(x = Inf, y = Inf, label = label),
              hjust = 1.1, vjust = 1.2, size = 4.5, color = "black", check_overlap = T) +
  facet_wrap( ~ symbol + cell_type, scales = "free", nrow = 2, ncol = 5) +
  # facet_wrap(~ celltype+symbol, scales = "free_y") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
      scale_color_manual(values = c(
      "Excitatory Neurons" = "#00008B",
      "Inhibitory Neurons" = "#CC3311",
      "Microglia" = "#FF8C00",
      "Astrocytes" = "#9932CC",
      "Oligodendrocytes" = "#698b22",
      "OPCs" = "#008b45",
      "Endothelial Cells" = "#8a6407"
    )) +
  xlab("APOE e4") + 
  ylab("Expression") +
  theme_classic() +
  theme(legend.position = "none")

(IL17D_plot = ggarrange(p_IL17D_PRS, p_IL17D_APOE, ncol = 2, nrow = 1))

cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I//PRS_IL17D_plot.pdf", width = 5, height = 3)
IL17D_plot
dev.off()
```

```{r}
cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I//PRS_p1.pdf", width = 4, height = 3)
p1
dev.off()

cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I//PRS_p2.pdf", width = 12, height = 6)
p2
dev.off()

cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I//PRS_p3.pdf", width = 12, height = 6)
p3
dev.off()

cairo_pdf("/pastel/projects/cytokines_Juliana/JNI_revision_I//PRS_p4.pdf", width = 15, height = 2.5)
p4
dev.off()
```

# Session info
```{r session_info}
sessionInfo()
```
