---
title: "Significant associations - Cytokines and phenotypic data - snRNA-seq"
author: |
 | Juliana Beker Godoy, MSc
 | Graduate Program in Bioinformatics
 | Professional and Technological Education Sector
 | Federal University of Parana (UFPR)
 | Curitiba, Brazil
date: "`r Sys.Date()`"
output:
 rmarkdown::html_document:
   theme: cerulean
   highlight: haddock
   code_folding: hide
   toc: true
   toc_float: true
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

> 424 unique samples | 
> Dorsolateral prefrontal cortex (DLFPC) | 
> ROSMAP cohort 

# Cell types

```{r clean, warning=FALSE, echo=FALSE, results='hide'}
#This command clean all variables. BE CAREFULL!!! 
rm(list = setdiff(ls(), lsf.str()))
```

```{r HelperFunctions, echo = FALSE}
# Function to create useful tables
createDT <- function(DF, caption="", scrollY=500){
  data <- DT::datatable(DF, caption=caption,
   extensions =  'Buttons',
   options = list( dom = 'Bfrtip', 
                    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'), 
                    scrollY = scrollY, scrollX=T, scrollCollapse = T, paging = F,  
                    columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
   ) 
   return(data)
}
```

```{r packages, warning=FALSE, echo=FALSE, message=FALSE}
library(dplyr)
library(rstatix)
library(readr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(readr)
library(tibble)
library(ggeasy)
library(ggpubr)
library(ggsignif)
#BiocManager::install("ComplexHeatmap")
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
options(encoding = "UTF-8")
```

```{r set_dir, warning=FALSE, message=FALSE, include=FALSE} 
work_dir <- "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC"
```

```{r source_reg_function, warning=FALSE, message=FALSE, include=FALSE} 
source("C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/run_regressions_v2_adjcov.R") # Our script for regressions: linear and logistic 
```

```{r exprData_upload, warning=FALSE, message=FALSE, include=FALSE} 
# RDS file
file_path <- file.path(work_dir, "snRNA_pseudoBulk_7majCellTypes.rds")
celltype_exp <- readRDS(file_path)
```

```{r phenoData_upload, warning=FALSE, message=FALSE, include=FALSE} 
file_path <- "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/pheno_SN.Rdata"
load(file_path)
dim(pheno_SN)
```

```{r cytokines_list_upload, warning=FALSE, message=FALSE, include=FALSE} 
# upload list of cytokines
file_path <- "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/list_cytokines/T015/list_cytokines_T015.txt"
cytokines <- read.delim(file_path, header = TRUE, check.names = FALSE, stringsAsFactors = FALSE)
cytokines <- subset(cytokines, select = -family) # I removed the 'family' column here because it had NA values
ensembls = cytokines$ensembl
```

```{r function_corr_sn, message=FALSE, warning=FALSE, results='hide', echo=FALSE} 
run_cytokines_analysis <- function(celltype_list, list_name, output_filename_pdf, width_pdf, height_pdf, output_filename_png, width_png, height_png, title) {
  exprData = celltype_list[[list_name]]
  
  # Filter the gene expression
  exprData_filt = exprData[rownames(exprData) %in% cytokines$ensembl, ]
  exprData_filt$ensembl = rownames(exprData_filt)
  exprData_cytokines = merge(exprData_filt, cytokines, by = "ensembl")
  rownames(exprData_cytokines) = exprData_cytokines$symbol
  exprData_cytokines$ensembl = NULL
  exprData_cytokines$symbol = NULL

  # AD traits 
  # The function adjusts by age, sex, and educ
  pheno_list = c(
    "cogng_demog_slope"="gaussian",
    "cogng_path_slope"="gaussian",
    "tangles_sqrt"="gaussian",
    "amyloid_sqrt"="gaussian",
    "gpath"="gaussian",
    #"tdp_cs_6reg"="gaussian",
    "ad_dementia_status"="binomial",
    "ci_status"="binomial"
  )

  data4linear_reg <- as.data.frame(t(exprData_cytokines))
  phenotype_dt <- pheno_SN[match(rownames(data4linear_reg), pheno_SN$projid), ]

  res_test = run_module_trait_association(data4linear_reg, phenotype_dt, pheno_list, covariates = c("age_death","msex", "educ"))

  # Heatmap 
  to_show = colnames(res_test$matrix_pvalue)
  transpose = T
  show_only_significant = T; signif_cutoff = c("***","**","*")

  matrix_rsquared = res_test$matrix_rsquared
  matrix_pvalue = res_test$matrix_pvalue # final matrix with the pvalues 
  
  # Reorder heatmap row names to paper 
  row_newOrder <- c("gpath","tangles_sqrt","amyloid_sqrt","ad_dementia_status","ci_status","cogng_demog_slope","cogng_path_slope")
  matrix_rsquared = matrix_rsquared[row_newOrder, ]
  matrix_pvalue = matrix_pvalue[row_newOrder, ]

  matrix_rsquared_to_plot = matrix_rsquared[,to_show]
  matrix_pvalue_to_plot = matrix_pvalue[,to_show]

  # Adjust P-values by each phenotype separately.
  adj_matrix_pvalue_to_plot = matrix_pvalue_to_plot
  for(i in 1:ncol(matrix_pvalue_to_plot)){
    adj_matrix_pvalue_to_plot[,i] = p.adjust(matrix_pvalue_to_plot[,i], method = "bonferroni")
  }
  
  adj_matrix_pvalue_to_plot.signif <- symnum(x = as.matrix(adj_matrix_pvalue_to_plot), corr = FALSE, na = FALSE,
                                             cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
                                             symbols = c("***", "**", "*", ".", " "))

  log_matrix_pvalue_to_plot = -log10(matrix_pvalue_to_plot)
  dimnames(log_matrix_pvalue_to_plot) = dimnames(log_matrix_pvalue_to_plot)

  if(show_only_significant){
    if(is.numeric(signif_cutoff)){
      to_keep = colSums(adj_matrix_pvalue_to_plot <= signif_cutoff) > 0
    }else{
      to_keep = rep(F,ncol(adj_matrix_pvalue_to_plot.signif))
      for(cut_i in signif_cutoff){
        to_keep = to_keep | colSums(adj_matrix_pvalue_to_plot.signif == cut_i) > 0 # change for the significance you want 
      }
    }
    log_matrix_pvalue_to_plot = log_matrix_pvalue_to_plot[,to_keep]
    adj_matrix_pvalue_to_plot.signif = adj_matrix_pvalue_to_plot.signif[,to_keep]
  }

  matrix_pvalue_to_plot_labels = formatC(log_matrix_pvalue_to_plot, format = "f", digits = 2)
  log_matrix_pvalue_to_plot_t = t(log_matrix_pvalue_to_plot)

  if(transpose){
    log_matrix_pvalue_to_plot_t = t(log_matrix_pvalue_to_plot_t)
    matrix_pvalue_to_plot_labels = t(matrix_pvalue_to_plot_labels)
    adj_matrix_pvalue_to_plot.signif = t(adj_matrix_pvalue_to_plot.signif)
  }

  # Colored by -log10(pvalue)
  # Numbers inside cell = -log10(pvalue): nominal

# New column names
new_column_names <- c(
  "gpath" = "Global AD burden",
  "tangles_sqrt" = "PHFtau tangle density",
  "amyloid_sqrt" = "Amyloid-β load",
  "ad_dementia_status" = "AD diagnosis",
  "ci_status" = "Mild cognitive impairment",
  "cogng_demog_slope" = "Cognitive decline",
  "cogng_path_slope" = "Resilience"
)

rownames(log_matrix_pvalue_to_plot_t) <- new_column_names[rownames(log_matrix_pvalue_to_plot_t)]

# Group colors
group_colors <- c("Pathology" = "#800074", "Cognition" = "#298c8c")

# Y lab annotation
group_ylab <- factor(
  c("Pathology","Pathology","Pathology","Cognition","Cognition","Cognition","Cognition"),
  levels = c("Pathology", "Cognition")
)

# Lateral annotation
row_anno <- rowAnnotation(
  Group = group_ylab,
  col = list(Group = group_colors),
  show_annotation_name = TRUE,
  annotation_name_side = "top",
  annotation_name_gp = gpar(fontsize = 9),
  width = unit(1, "mm")
)

# Heatmap
heatmap_ext <- Heatmap(
  log_matrix_pvalue_to_plot_t,
  name = "-log10(P-value)",
  cell_fun = function(j, i, x, y, width, height, fill) {
    if (as.character(t(adj_matrix_pvalue_to_plot.signif)[i, j]) == " ") {
      grid.text(t(matrix_pvalue_to_plot_labels)[i, j], x, y, gp = gpar(fontsize = 8))
    } else {
      grid.text(
        paste0(
          t(matrix_pvalue_to_plot_labels)[i, j], "\n",
          t(adj_matrix_pvalue_to_plot.signif)[i, j]
        ),
        x, y, gp = gpar(fontsize = 8)
      )
    }
  },
  col = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
  row_names_side = "right",
  show_row_names = TRUE,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  column_names_gp = gpar(fontsize = 9, fontface = "italic"),
  row_names_gp = gpar(fontsize = 9),  
  border = TRUE,
  show_row_dend = FALSE,
  show_column_dend = FALSE,
  rect_gp = gpar(col = "white", lwd = 1),
  column_names_rot = 40,
  column_title = title
)

final_heatmap <- row_anno + heatmap_ext

ht_draw <- draw(final_heatmap, padding = unit(c(3, 4, 3, 3), "mm"))  # Ajustar as margens (top, right, bottom, left)

# Saving as PDF
pdf(file = output_filename_pdf, width = width_pdf, height = height_pdf)
draw(final_heatmap, padding = unit(c(5, 4, 3, 4), "mm"))
dev.off()

# Saving as PNG
png(file = output_filename_png, width = width_png, height = height_png, res = 300)
draw(final_heatmap, padding = unit(c(5, 4, 3, 4), "mm"))
dev.off()
  
  # Return the result object
   return(list(
    res_test = res_test,
    adj_matrix_pvalue_to_plot = adj_matrix_pvalue_to_plot,
    matrix_rsquared = res_test$matrix_rsquared,
    signif = matrix_pvalue_to_plot_labels
  ))
}
```

## Excitatory Neurons

```{r ext_neurons_reg, message=FALSE, results='hide', warning=FALSE, echo=FALSE, fig.width=5.9, fig.height=3.2} 
# Run analysis
result_ext <- run_cytokines_analysis(celltype_exp, "ext",
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_ext_T043.pdf", width_pdf = 5.5, height_pdf = 2.5,
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_ext_T043.png", width_png = 1800, height_png = 875, title = "Excitatory Neurons")
```

### Rsquared

```{r squared_ext}
createDT(result_ext$matrix_rsquared)
```

### Adjusted pvalue

```{r pvalue_ext}
createDT(result_ext$adj_matrix_pvalue_to_plot)
```

## Inhibitory Neurons

```{r inh_neurons_reg, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=5, fig.height=3.5} 
result_inh <- run_cytokines_analysis(celltype_exp, "inh",
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_inh_T043.pdf", width_pdf = 4.60, height_pdf = 3.10,
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_inh_T043.png", width_png = 1500, height_png = 900, title = "Inhibitory Neurons")
```

### Rsquared

```{r squared_inh}
createDT(result_inh$matrix_rsquared)
```

### Adjusted pvalue

```{r pvalue_inh}
createDT(result_inh$adj_matrix_pvalue_to_plot)
```

## Microglia

```{r mic_reg, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=5.5, fig.height=3} 

result_mic <- run_cytokines_analysis(celltype_exp, "mic",
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_mic_T043.pdf", width_pdf = 5.5, height_pdf = 3,
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_mic_T043.png", width_png = 1720, height_png = 940, title = "Microglia")
```


### Rsquared

```{r squared_mic}
createDT(result_mic$matrix_rsquared)
```

### Adjusted pvalue

```{r pvalue_mic}
createDT(result_mic$adj_matrix_pvalue_to_plot)
```

## Oligodendrocytes

```{r oli_reg, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=4.6, fig.height=3.5} 

result_oli <- run_cytokines_analysis(celltype_exp, "oli",
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_oli_T043.pdf", width_pdf = 4.25, height_pdf = 3,
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_oli_T043.png", width_png = 1410, height_png = 825, title = "Oligodendrocytes")
```

### Rsquared

```{r squared_oli}
createDT(result_oli$matrix_rsquared)
```

### Adjusted pvalue

```{r pvalue_oli}
createDT(result_oli$adj_matrix_pvalue_to_plot)
```

## Astrocytes

```{r ast_reg, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=6, fig.height=3.25} 

result_ast <- run_cytokines_analysis(celltype_exp, "ast",
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_ast_T043.pdf", width_pdf = 6, height_pdf = 3.25,
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_ast_T043.png", width_png = 1800, height_png = 900, title = "Astrocytes")
```

### Rsquared

```{r squared_ast}
createDT(result_ast$matrix_rsquared)
```

### Adjusted pvalue

```{r pvalue_ast}
createDT(result_ast$adj_matrix_pvalue_to_plot)
```

## Oligodendrocyte progenitor cells

```{r opc_reg, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.width=5, fig.height=3.25} 

result_opc <- run_cytokines_analysis(celltype_exp, "opc",
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_opc_T043.pdf", width_pdf = 4.8, height_pdf = 3.25,
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_opc_T043.png", width_png = 1475, height_png = 950, title = "OPCs")
```

### Rsquared

```{r squared_opc}
createDT(result_opc$matrix_rsquared)
```

### Adjusted pvalue

```{r pvalue_opc}
createDT(result_opc$adj_matrix_pvalue_to_plot)
```

## Endothelial cells

There is no significant results to this cell type

```{r end_reg, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=5, fig.height=6, include=FALSE} 

result_end <- run_cytokines_analysis(celltype_exp, "end",
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_end_T043.pdf", width_pdf = 7, height_pdf = 6,
                                     "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/reg_signif_end_T043.png", width_png = 1500, height_png = 1200, title = "Endothelial cells")
```

### Nominal pvalue

```{r pvalue_end}
createDT(result_end$adj_matrix_pvalue_to_plot)
```

```{r list_signif_genes, message=FALSE, warning=FALSE, include=FALSE}
# Making a list with expressed genes in sn data

# Ext
list_signif_ext <- rownames(result_ext$signif)
save(list_signif_ext, file = "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/list_signif_ext_T043.RData")

# Inh
list_signif_inh <- rownames(result_inh$signif)
save(list_signif_inh, file = "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/list_signif_inh_T043.RData") 

# Mic
list_signif_mic <- rownames(result_mic$signif)
save(list_signif_mic, file = "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/list_signif_mic_T043.RData") 

# Ast
list_signif_ast <- rownames(result_ast$signif)
save(list_signif_ast, file = "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/list_signif_ast_T043.RData") 

# Oli
list_signif_oli <- rownames(result_oli$signif)
save(list_signif_oli, file = "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/list_signif_oli_T043.RData") 

# Opc
list_signif_opc <- rownames(result_opc$signif)
save(list_signif_opc, file = "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/T043/list_signif_opc_T043.RData") 
```

# Session info
```{r}
sessionInfo()
```