---
title: "Significant associations - Data from T019 - snRNA-seq"
author: |
 | Ricardo A. Vialle, Rush Alzheimer's Disease Center
 | Juliana Beker Godoy, MSc - Graduate program in Bioinformatics, UFPR, Curitiba, Brazil
date: "`r Sys.Date()`"
output:
 rmarkdown::html_document:
   theme: cerulean
   highlight: haddock
   code_folding: hide
   toc: true
   toc_float: true
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

# Linear/Logistic Regression

**Test:** [T043](https:/docs.google.com/spreadsheets/d/11Ts8o1gQrzWQ4KoI7_UuudokQ-LxWyYY8iEbXD6sdyU/edit#gid=0&range=A45)

**Date:** 4/9/2024

**Link Github:** https:/github.com/RushAlz/Cytokines/tree/main/sn_RNAseq_DLPFC/T043

**Objective:** Filter to show only significant results from the linear and/or logistic regressions from single-nuclei data (T019)

```{r clean, warning=FALSE, echo=FALSE, results='hide'}
#This command clean all variables. BE CAREFULL!!! 
rm(list = setdiff(ls(), lsf.str()))
```

```{r HelperFunctions, echo = FALSE}
# Function to create useful tables
createDT <- function(DF, caption="", scrollY=500){
  data <- DT::datatable(DF, caption=caption,
   extensions =  'Buttons',
   options = list( dom = 'Bfrtip', 
                    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'), 
                    scrollY = scrollY, scrollX=T, scrollCollapse = T, paging = F,  
                    columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
   ) 
   return(data)
}
```

```{r packages, warning=FALSE, echo=FALSE, message=FALSE}
library(dplyr)
library(rstatix)
library(readr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(readr)
library(tibble)
library(ggeasy)
library(ggpubr)
library(ggsignif)
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
options(encoding = "UTF-8")
```

```{r set_dir, warning=FALSE, message=FALSE, include=FALSE} 
# work_dir <- "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC"
work_dir <- "/pastel/projects/cytokines_Juliana/JNI_revision_I/"
```

```{r source_reg_function, warning=FALSE, message=FALSE, include=FALSE} 
source("/pastel/projects/cytokines_Juliana/JNI_revision_I/run_regressions_v3_adjcov.R") # Our script for regressions: linear and logistic 
```

## Read data and prepare inputs

```{r exprData_upload, warning=FALSE, message=FALSE, include=FALSE} 
# load("/pastel/resources/data2share/PMI.RData")

syn_idmap = read_tsv("/pastel/resources/synapse/SynapseID_mapping.tsv", show_col_types = F)
batch_metadata = readxl::read_xlsx("/pastel/resources/20220203_snRNAseq_AMPAD/41586_2024_7871_MOESM5_ESM.xlsx", 
                           sheet = "Participant inclusion QCs", col_names = TRUE)
batch_metadata = batch_metadata %>% left_join(syn_idmap, by = c("individualID" = "synapseid")) # Join to get projid
batch_metadata = batch_metadata %>% select(projid, batch, min_LOD_score)

# RIN = read_csv("/pastel/resources/20220203_snRNAseq_AMPAD/RIN_n1089_08222023.csv")
# RIN$projid = sprintf("%08d", RIN$projid) # Ensure projid is in the correct format
# sum(RIN$projid %in% batch_metadata$projid) # 410

# Caminho completo para o arquivo RDS
file_path <- file.path(work_dir, "snRNA_pseudoBulk_7majCellTypes.rds")

# Carregue o arquivo RDS
celltype_exp <- readRDS(file_path)
lapply(celltype_exp, dim) # Check dimensions of each cell type
```

```{r phenoData_upload, warning=FALSE, message=FALSE, include=FALSE} 
# file_path <- "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/sn_RNAseq_DLPFC/pheno_SN.Rdata"
# file_path <- paste0(work_dir, "pheno_SN.Rdata")
# load(file_path)

# # Prepare new phenotypes
# pheno1 = as.data.frame(readRDS("/pastel/resources/data2share/basic_Apr2022_customJul2024.rds"))
# pheno2 = as.data.frame(readRDS("/pastel/resources/data2share/basic_Apr2022_selected_list_Jul2024.rds"))
# 
# celltype_exp <- readRDS(file.path(work_dir, "snRNA_pseudoBulk_7majCellTypes.ComBat.rds"))
# load("/pastel/resources/data2share/bulk_MF/bulk_DLPFC_2022.Rdata")
# celltype_exp[["bulk DLPFC"]] = exprData_DLPFC
# 
# all_donors = lapply(celltype_exp, function(x) colnames(x)) %>% unlist() %>% unique()
# 
# selected_variables = c("projid","pmi","cogdx_3grp",colnames(pheno2))
# 
# new_phenotypes = pheno1[pheno1$projid %in% all_donors, selected_variables] %>% distinct()
# saveRDS(new_phenotypes, file.path(work_dir, "phenotypes_updated.rds"))

pheno_SN = as.data.frame(readRDS(file.path(work_dir, "phenotypes_updated.rds")))
rownames(pheno_SN) = pheno_SN$projid
pheno_SN = pheno_SN %>% 
  # left_join(pmi_df) %>% 
  # left_join(RIN) %>%
  left_join(batch_metadata)

# Replace NA values in 'pmi' with the mean of non-NA values
pheno_SN$pmi[is.na(pheno_SN$pmi)] <- mean(pheno_SN$pmi, na.rm = TRUE)

dim(pheno_SN)
```

```{r cytokines_list_upload, warning=FALSE, message=FALSE, include=FALSE} 
# upload list of cytokines
# file_path <- "C:/Users/beker/OneDrive/Documentos/Mestrado/GitHub/Cytokines/list_cytokines/T015/list_cytokines_T015.txt"
file_path <- paste0(work_dir, "list_cytokines_T015.txt")
cytokines <- read.delim(file_path, header = TRUE, check.names = FALSE, stringsAsFactors = FALSE)
cytokines <- subset(cytokines, select = -family) # I removed the 'family' column here because it had NA values
ensembls = cytokines$ensembl

# genes = list() 
# for(cell_i in names(celltype_exp)) {
#   exprData = celltype_exp[[cell_i]]
#   exprData_filt = exprData[rownames(exprData) %in% cytokines$ensembl, ]
#   genes[[cell_i]] = rownames(exprData_filt)
# }
# lapply(genes, length) # Check how many genes are in each cell type
# n_tests = 7 * sum(unlist(lapply(genes, length))) # Total number of genes across all cell types
# # 160 genes in total * 7 pheno types = 1120 tests
```

```{r eval=F}
library(limma)
library(edgeR)

norm_method = "none"

# Define the covariates for the model
covs <- c("batch","pmi")

for(cell_i in names(celltype_exp)) {
  exprData = celltype_exp[[cell_i]]
  exprData.meta = pheno_SN[match(colnames(exprData), pheno_SN$projid), ]
  
  # Construct the design matrix from covariates
  design_formula <- paste0("~", paste0(covs, collapse = "+")) %>% as.formula()
  design <- model.matrix(design_formula, data = exprData.meta)
  design <- design[, colSums(design) > 0] # Remove columns with zero sum

  # Normalize counts and fit linear model using voom and lmFit
  fit <- lmFit(exprData, design) 
  
  # Adjust data based on fitted model, excluding the intercept
  indx <- grepl("(Intercept)", colnames(fit$design))
  adjusted_data <- exprData - (fit$design[, !indx] %*% t(fit$coefficients[, !indx]))
  celltype_exp[[cell_i]] = adjusted_data
}
```

```{r eval=F}
library(sva)
for(cell_i in names(celltype_exp)) {
  exprData = celltype_exp[[cell_i]]
  exprData.meta = pheno_SN[match(colnames(exprData), pheno_SN$projid), ]
  
  adjusted_data <- ComBat(exprData, exprData.meta$batch)
  celltype_exp[[cell_i]] = adjusted_data
}
saveRDS(celltype_exp, file.path(work_dir, "snRNA_pseudoBulk_7majCellTypes.ComBat.rds"))
```

## Prepare parameters

```{r}
celltype_exp <- readRDS(file.path(work_dir, "snRNA_pseudoBulk_7majCellTypes.ComBat.rds"))
load("/pastel/resources/data2share/bulk_MF/bulk_DLPFC_2022.Rdata")

celltype_exp[["bulk DLPFC"]] = exprData_DLPFC

covariates = c("age_death","msex", "educ", "pmi")

pheno_list = c(
  "cogng_demog_slope"="gaussian",
  "cogng_path_slope"="gaussian",
  "tangles_sqrt"="gaussian",
  "amyloid_sqrt"="gaussian",
  "gpath"="gaussian",
  "ad_dementia_status"="binomial",
  "cog_impairment_status"="binomial"
)
```

## Run association tests

```{r eval=T}
assoc_results = list()
for(cell_i in names(celltype_exp)) {
  exprData = as.data.frame(celltype_exp[[cell_i]])
  
  # Filter the gene expression
  exprData_ensembl = gsub("(.*)\\.(.*)","\\1", rownames(exprData)) 
  exprData_filt = exprData[exprData_ensembl %in% cytokines$ensembl, ]
  exprData_filt$ensembl = gsub("(.*)\\.(.*)","\\1",rownames(exprData_filt))
  exprData_cytokines = merge(exprData_filt, cytokines, by = "ensembl")
  rownames(exprData_cytokines) = exprData_cytokines$symbol
  exprData_cytokines$ensembl = NULL
  exprData_cytokines$symbol = NULL
  
  data4linear_reg <- as.data.frame(t(exprData_cytokines))
  phenotype_dt <- pheno_SN[match(rownames(data4linear_reg), pheno_SN$projid), ]

  res_test = run_module_trait_association(data4linear_reg, 
                                          phenotype_dt, 
                                          pheno_list,
                                          covariates = covariates)
  
  assoc_results[[cell_i]] = res_test
}
save(assoc_results, file = file.path(work_dir, "cytokines_snRNA_association_results.RData"))
```

```{r}
load(file.path(work_dir, "cytokines_snRNA_association_results.RData"))
```

## bulk DLPFC

```{r, message=FALSE, results='hide', warning=FALSE, echo=FALSE, fig.width=8, fig.height=10}
source("/pastel/projects/cytokines_Juliana/JNI_revision_I/run_regressions_v3_adjcov.R") # Our script for regressions: linear and logistic 
result_plot <- create_combined_visualization(assoc_results$`bulk DLPFC`, "DLPFC bulk RNAseq")

png(paste0(work_dir,"/reg_signif_bulk_T043_review1.png"), width = 8, height = 10, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_bulk_T043_review1.pdf"), width = 8, height = 10)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$`bulk DLPFC`$all_stats_df)
```

## bulk DLPFC (male only)

```{r, message=FALSE, results='hide', warning=FALSE, echo=FALSE, fig.width=8, fig.height=5}
female_ids = pheno_SN %>% filter(msex == 0) %>% filter(projid %in% colnames(exprData_DLPFC)) %>% pull(projid)
celltype_exp[["bulk DLPFC (female)"]] = exprData_DLPFC[,female_ids]

male_ids = pheno_SN %>% filter(msex == 1) %>% filter(projid %in% colnames(exprData_DLPFC)) %>% pull(projid)
exprData = exprData_DLPFC[,male_ids]

covariates = c("age_death","educ", "pmi")

# Filter the gene expression
exprData_ensembl = gsub("(.*)\\.(.*)","\\1", rownames(exprData)) 
exprData_filt = exprData[exprData_ensembl %in% cytokines$ensembl, ]
exprData_filt$ensembl = gsub("(.*)\\.(.*)","\\1",rownames(exprData_filt))
exprData_cytokines = merge(exprData_filt, cytokines, by = "ensembl")
rownames(exprData_cytokines) = exprData_cytokines$symbol
exprData_cytokines$ensembl = NULL
exprData_cytokines$symbol = NULL

data4linear_reg <- as.data.frame(t(exprData_cytokines))
phenotype_dt <- pheno_SN[match(rownames(data4linear_reg), pheno_SN$projid), ]

res_test = run_module_trait_association(data4linear_reg, 
                                        phenotype_dt, 
                                        pheno_list,
                                        covariates = covariates)
  
  
result_plot <- create_combined_visualization(res_test, "DLPFC bulk RNAseq (males only)")

png(paste0(work_dir,"/reg_signif_bulk_males_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_bulk_males_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(res_test$all_stats_df)
```

## bulk DLPFC (female only)

```{r, message=FALSE, results='hide', warning=FALSE, echo=FALSE, fig.width=8, fig.height=10}
female_ids = pheno_SN %>% filter(msex == 0) %>% filter(projid %in% colnames(exprData_DLPFC)) %>% pull(projid)
exprData = exprData_DLPFC[,female_ids]

covariates = c("age_death","educ", "pmi")

# Filter the gene expression
exprData_ensembl = gsub("(.*)\\.(.*)","\\1", rownames(exprData)) 
exprData_filt = exprData[exprData_ensembl %in% cytokines$ensembl, ]
exprData_filt$ensembl = gsub("(.*)\\.(.*)","\\1",rownames(exprData_filt))
exprData_cytokines = merge(exprData_filt, cytokines, by = "ensembl")
rownames(exprData_cytokines) = exprData_cytokines$symbol
exprData_cytokines$ensembl = NULL
exprData_cytokines$symbol = NULL

data4linear_reg <- as.data.frame(t(exprData_cytokines))
phenotype_dt <- pheno_SN[match(rownames(data4linear_reg), pheno_SN$projid), ]

res_test = run_module_trait_association(data4linear_reg, 
                                        phenotype_dt, 
                                        pheno_list,
                                        covariates = covariates)
  
  
result_plot <- create_combined_visualization(res_test, "DLPFC bulk RNAseq (females only)")

png(paste0(work_dir,"/reg_signif_bulk_females_review1.png"), width = 8, height = 10, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_bulk_females_review1.pdf"), width = 8, height = 10)
result_plot$plot
dev.off()
```

```{r }
createDT(res_test$all_stats_df)
```

## Excitatory Neurons

```{r, message=FALSE, results='hide', warning=FALSE, echo=FALSE, fig.width=8, fig.height=5}
result_plot <- create_combined_visualization(assoc_results$ext, "Excitatory Neurons")

png(paste0(work_dir,"/reg_signif_ext_T043_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_ext_T043_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$ext$all_stats_df)
```

## Inhibitory Neurons

```{r, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=8, fig.height=5} 
result_plot <- create_combined_visualization(assoc_results$inh, "Inhibitory Neurons")

png(paste0(work_dir,"/reg_signif_inh_T043_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_inh_T043_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$inh$all_stats_df)
```

## Microglia

```{r, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=8, fig.height=5} 
result_plot <- create_combined_visualization(assoc_results$mic, "Microglia")

png(paste0(work_dir,"/reg_signif_mic_T043_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_mic_T043_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$mic$all_stats_df)
```

## Oligodendrocytes

```{r, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=8, fig.height=5} 
result_plot <- create_combined_visualization(assoc_results$oli, "Oligodendrocytes")

png(paste0(work_dir,"/reg_signif_oli_T043_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_oli_T043_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$oli$all_stats_df)
```

## Astrocytes

```{r, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=8, fig.height=5} 
result_plot <- create_combined_visualization(assoc_results$ast, "Astrocytes")

png(paste0(work_dir,"/reg_signif_ast_T043_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_ast_T043_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$ast$all_stats_df)
```

## Oligodendrocyte progenitor cells

```{r, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=8, fig.height=5} 
result_plot <- create_combined_visualization(assoc_results$opc, "OPCs")

png(paste0(work_dir,"/reg_signif_opc_T043_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_opc_T043_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$opc$all_stats_df)
```

## Endothelial cells

```{r, message=FALSE, warning=FALSE, results='hide', echo=FALSE, fig.width=8.5, fig.height=5} 
result_plot <- create_combined_visualization(assoc_results$end, "Endothelial cells")

png(paste0(work_dir,"/reg_signif_end_T043_review1.png"), width = 8, height = 5, res = 600, units = "in")
result_plot$plot
dev.off()
cairo_pdf(paste0(work_dir,"/reg_signif_end_T043_review1.pdf"), width = 8, height = 5)
result_plot$plot
dev.off()
```

```{r }
createDT(assoc_results$end$all_stats_df)
```

## IL-15

```{r, message=FALSE, warning=FALSE} 
celltype_exp <- readRDS(file.path(work_dir, "snRNA_pseudoBulk_7majCellTypes.ComBat.rds"))
load("/pastel/resources/data2share/bulk_MF/bulk_DLPFC_2022.Rdata")

celltype_exp[["bulk DLPFC"]] = exprData_DLPFC

dt_res1 = as.data.frame(celltype_exp[["mic"]], check.names = F) 
dt_res1_selected = dt_res1[rownames(dt_res1) %in% cytokines$ensembl, ]
# long format for the plot
dt_res1_selected_l = dt_res1_selected %>% 
  rownames_to_column("ensembl") %>%
  pivot_longer(-ensembl, values_to = "expression", names_to = "projid") %>%
  mutate(celltype = "mic")

dt_res2 = as.data.frame(celltype_exp[["bulk DLPFC"]], check.names = F) 
dt_res2_selected = dt_res2[gsub("(.*?)\\.(.*)","\\1",rownames(dt_res2)) %in% cytokines$ensembl, ]
# long format for the plot
dt_res2_selected_l = dt_res2_selected %>% 
  rownames_to_column("ensembl") %>%
  mutate(ensembl = gsub("(.*?)\\.(.*)","\\1",ensembl)) %>%
  pivot_longer(-ensembl, values_to = "expression", names_to = "projid") %>%
  mutate(celltype = "bulk DLPFC")

# join to get the gene_name
dt_res_selected_l = bind_rows(dt_res1_selected_l,dt_res2_selected_l) %>%
  left_join(cytokines, by = "ensembl")
```

### IL15 expression by diagnosis

```{r, message=FALSE, warning=FALSE, fig.width=4.5, fig.height=3.5} 
# join to get the phenotype 
dt_res_selected_l_meta = dt_res_selected_l %>% 
  left_join(pheno_SN, by = "projid") %>%
  filter(symbol == "IL15")

# By diagnosis 
dt_res_selected_l_meta$cogdx_3grp[dt_res_selected_l_meta$cogdx_3grp == 0] <- "NCI"
dt_res_selected_l_meta$cogdx_3grp[dt_res_selected_l_meta$cogdx_3grp == 1] <- "MCI"
dt_res_selected_l_meta$cogdx_3grp[dt_res_selected_l_meta$cogdx_3grp == 2] <- "AD"

dt_res_selected_l_meta$cogdx_3grp = factor(dt_res_selected_l_meta$cogdx_3grp, levels = c("NCI", "MCI", "AD"))
names(dt_res_selected_l_meta)[names(dt_res_selected_l_meta) == 'cogdx_3grp'] <- 'Diagnosis'

p <- ggplot(na.omit(dt_res_selected_l_meta[, c("Diagnosis", "expression", "symbol", "celltype")]), aes(x = Diagnosis, y = expression, color = Diagnosis)) + 
  geom_boxplot(notch=F, fill="white", color="black", outlier.shape=NA, lwd=0.4, alpha=1, width=.7, fatten = 3) +
  ggbeeswarm::geom_quasirandom(alpha=.8, width=.2, show.legend=F, size=2) + 
  geom_pwc(method = "t_test", label = "p.adj.signif", p.adjust.method = "bonferroni",
           label.y.npc = "top", label.size = 5, hide.ns = TRUE) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
                   geom = "crossbar", width = .7, color = "black") +
  facet_wrap(~celltype, scales = "free_y", nrow = 1, strip.position = "left", 
                labeller = as_labeller(c(`bulk DLPFC` = "IL15 expression in bulk DLPFC", mic = "IL15 expression in microglia") ) ) +
  # facet_wrap(~ celltype+symbol, scales = "free_y") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_color_manual(values = c("AD" = "#BB5566", "MCI" = "#DDAA33", "NCI" = "#004488")) + 
  xlab("Final consensus diagnosis") + 
  ylab(NULL) +
  labs(title = "IL15 expression by diagnosis") +
  theme_classic() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 11),
        strip.placement = "outside",
        legend.position = "none")

# Print the plot
print(p)
  
# Save to PDF
cairo_pdf(paste0(work_dir,"/boxplot_IL15_by_diag_review1.pdf"), width = 4.5, height = 3.5)
print(p)
dev.off()

# Save to PNG
png(paste0(work_dir,"/boxplot_IL15_by_diag_review1.png"), width = 4.5, height = 3.5, res = 600, units = "in")
print(p)
dev.off()
```

### IL15 expression by gpath (mic)

```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=3.5} 
# By gpath
p = ggplot(na.omit(dt_res_selected_l_meta[dt_res_selected_l_meta$celltype == "mic",
                                      c("gpath","Diagnosis", "expression", "symbol", "celltype")]), 
       aes(x = expression, y = gpath, color = Diagnosis)) + 
  geom_point(size = 1.5, alpha = 0.6) +
  stat_smooth(method = "lm", se = F) + # Add regression line
  stat_cor(method = "spearman", label.x.npc = "left", label.y.npc = "top", size = 4, show.legend = F) +
  easy_labs(y = "Global AD burden", x = "IL15 expression", title = "IL15 expression on microglia for global AD burden") +
  scale_color_manual(values = c("AD" = "#BB5566", "MCI" = "#DDAA33", "NCI" = "#004488")) + 
  theme_classic()

# Print the plot
print(p)
  
# Save to PDF
cairo_pdf(paste0(work_dir,"/scatterplot_IL15_by_gpath_mic_review1.pdf"), width = 5, height = 3.5)
print(p)
dev.off()

# Save to PNG
png(paste0(work_dir,"/scatterplot_IL15_by_gpath_mic_review1.png"), width = 5, height = 3.5, res = 600, units = "in")
print(p)
dev.off()
```

### IL15 expression by cog. decline (bulk)

```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=3.5} 
# By cogng_demog_slope
p = ggplot(na.omit(dt_res_selected_l_meta[dt_res_selected_l_meta$celltype == "bulk DLPFC",
                                      c("cogng_demog_slope","Diagnosis", "expression", "symbol", "celltype")]), 
       aes(x = expression, y = cogng_demog_slope, color = Diagnosis)) + 
  geom_point(size = 1.5, alpha = 0.5) +
  stat_smooth(method = "lm", se = F) + # Add regression line
  stat_cor(method = "spearman", label.x.npc = "left", label.y.npc = 0.2, size = 4, show.legend = F) +
  easy_labs(y = "Cognitive decline (slope)", x = "IL15 expression", title = "IL15 expression on bulk DLPFC for cognitive decline") +
  scale_color_manual(values = c("AD" = "#BB5566", "MCI" = "#DDAA33", "NCI" = "#004488")) + 
  theme_classic()

# Print the plot
print(p)
  
# Save to PDF
cairo_pdf(paste0(work_dir,"/scatterplot_IL15_by_cogdec_bulk_review1.pdf"), width = 5, height = 3.5)
print(p)
dev.off()

# Save to PNG
png(paste0(work_dir,"/scatterplot_IL15_by_cogdec_bulk_review1.png"), width = 5, height = 3.5, res = 600, units = "in")
print(p)
dev.off()
```

# Session info
```{r}
sessionInfo()
```
